Bayesian exponential random graph modeling of KWL Burial network

### 2020-12-04 analysis/paper/paper.Rmd
work on discussion section and add references

<!-- archaeological implication based on the results-->
The notable finding in our results is the change in network properties of Kiwulan burials from a more cohesive network (represented by higher transitivity and negative popularity) toward a more centralized network (lower transitivity and positive popularity) after the European presence during the 17th century. This confirms out hypothesized models of chronological differences in the structure of burials based on grave goods and burial attributes associated with demography and mortuary practices. The differences are in line with the theoretical assumption derived from the corporate-network continuum for explaining horizontal hierarchy that lacks characteristics of the traditional definition of vertical hierarchy but shows some level of social inequality [@Feinman2000]. This theoretical framework works better to understand the small scale societies and the strategies for gaining power between ambitious individuals in the pericolonical context.

<!-- methodological implication-->
The ERGM network analysis within a Bayesian framework provides a robust way to examine relations between burials based on grave goods and attributes associated with demography and mortuary practices. The adoption of Bayesian inference on ERGM can solve the issues of computational tractability and degeneracy in parameter estimation in ERGM [@Caimo2017]. In addition, a Bayesian framework reduced the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations. We demonstrated a plausible and robust method to test the hypothesized model of chronological changes in burial structures with probabilistic interpretation. It is important to note that although some network statistics of nodal attributes do not highly match our assumption, the Bayesian approach is still an efficient and effective method for evaluating the endogenous network effects. This closely related to the overall structure of burials that can be viewed as a representation of a given society.

>>>>

<!-- theoretical framework for the results-->
The notable finding in our results is the change in network properties of Kiwulan burials from a more cohesive network (represented by higher transitivity and negative popularity) toward a more centralized network (lower transitivity and positive popularity) after the European presence during the 17th century. This confirms our hypothesized models of chronological differences in the structure of burials in terms of grave goods and burial attributes associated with demography and mortuary practices. The differences are in line with theoretical assumptions of horizontal hierarchy, the corporate-network continuum, that focuses on distinct strategies for achieving power as an alternative explanation to understand the emergence and level of social inequality [@Drennan2010; @Feinman2000]. Compared to vertical hierarchy that stresses stratified social characteristics, the corporate-network model is more effective to investigate the way how a small scale society organized and explore changes in social relation in a pericolonial context. A corporate-based society stresses shared power between individuals, integrative rituals, and social inequality, if any, would be associated with a social group or sector [@Siegel1999]. In contrast, a network-based society presents wealth accumulation through individual networks, resources control, and trade monopolization.

<!--archaeological implication based on the results-->
The covariants for both networks before and after the European presence provide further information of which variables relative to the formation of networks with the same prestige goods in common. We included age, sex, ritual pottery, and wealth index based on value of grave goods. The pre-European network indicates that burials with pottery as ritual vessel tend to be tied, while the burials within the same level of wealth tend to be connected in the post-European network. Moreover, the burials with ritual vessel tend not to be tied in the post-European network. This illustrates there are different mechanisms to form relations between burials in terms of the access to trade trade goods. Before the European arrival, ritual behavior seem to be an emphasized cultural practice relative to the use of prestige goods. After the European arrival, the cultural emphasis shifted from ritual to wealth difference. The demographic information, such as age and sex, do not show any correlations with the network formation for both periods. This may be influenced by some unidentified burials that limits a more insightful result. The distance between burials is not a variable that correlates to the network formation either.

<!-- fit into a big anthropological picture-->

<!-- methodological implication-->
In this paper, we used burial network as a proxy to explore the social relations based on the argument that burials could be viewed as a social representation [@Binford1971; @Saxe1970]. We also note that the issue of disconnection between the role of the deceased in burial context and their real life, which can be resolved by examined different evidence out of the burial context [@Chapman2003]. The ERGM network analysis within a Bayesian framework provides a robust way to examine relations between burials based on grave goods and attributes associated with demography and mortuary practices. The adoption of Bayesian inference on ERGM can solve the issues of computational tractability and degeneracy in parameter estimation in ERGM [@Caimo2017]. In addition, a Bayesian framework reduced the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations. We demonstrated a plausible and robust method to test the hypothesized model of chronological changes in burial structures with probabilistic interpretation. It is important to note that although some network statistics of nodal attributes do not highly match our assumption, the Bayesian approach is still an efficient and effective method for evaluating the endogenous network effects. This closely related to the overall structure of burials that can be viewed as a representation of a given society.
